---
title: "Analyzing Gapminder Data"
author: "Kathleen Tran/Eric Almeida"
date: "3/6/2017"
output: pdf_document
---


```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(curl)
library(dplyr)
library(tidyr)

gapminder_location <-"https://raw.githubusercontent.com/resbaz/r-novice-gapminder-files/master/data/gapminder-FiveYearData.csv"

gapminder_location<-curl(url = "https://raw.githubusercontent.com/resbaz/r-novice-gapminder-files/master/data/gapminder-FiveYearData.csv")

gapminder<-read.csv(gapminder_location)
```

###We wanted to look at how life expectancy played a part in GDP per Capita, not only via all the continents available but also comparing two continents to see similarities, differences, and variations. Different types of graphs were created to see if trends existed.

```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE, fig.width =7, fig.height=7}
ggplot(gapminder, aes(x=lifeExp, y=gdpPercap, color=continent))+geom_point()+scale_y_log10()+geom_smooth(method=lm)+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")+ggtitle("The Life Expectancy of 5 continents via GDP per Capita")+labs(caption="(The R^2 value is 0.3403 and the p value is 2.2x10^-16.)")

ggplot(subset(gapminder, continent %in% c("Asia", "Africa")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_point()+geom_smooth()+ggtitle("Asia and Africa")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")

ggplot(subset(gapminder, continent %in% c("Asia", "Africa")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_bar(stat="identity")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")

ggplot(subset(gapminder, continent %in% c("Americas", "Europe")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_point()+ggtitle("Americas and Europe")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")

ggplot(subset(gapminder, continent %in% c("Americas", "Europe")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_bar(stat="identity")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")

ggplot(subset(gapminder, continent %in% c("Oceania", "Asia")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_point()+ggtitle("Oceania and Asia")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")

ggplot(subset(gapminder, continent %in% c("Oceania", "Asia")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_bar(stat="identity")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")
```

###We wanted to explore each continent and its average population, average GDP per Capita, and average life expectancy at a specific year. We chose to look at 2007 because it is the most recent year with available data.

```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE, fig.width =7, fig.height=7}
gPop <- group_by(gapminder, continent)
summarise(gPop,Population=mean(pop))
gapminder %>% group_by(continent) %>% filter (year==2007) %>% summarise(Population=mean(pop))%>% ggplot(aes(x=continent,y=Population, color=continent, fill=continent))+geom_bar(stat="identity")+scale_x_discrete(name="Continent")+scale_y_discrete(name="MeanPopulation")

gGDP <- group_by(gapminder, continent)
summarise(gGDP, GDP=mean(gdpPercap))
gapminder %>% group_by(continent) %>% filter (year==2007) %>% summarise(GDP=mean(gdpPercap)) %>% ggplot(aes(x=continent,y=GDP, color=continent, fill=continent))+geom_bar(stat="identity")+scale_x_discrete(name="Continent")+scale_y_discrete(name="MeanGDP")

gExp <- group_by(gapminder, continent)
summarise(gExp,LifeExpectancy=mean(lifeExp))
gapminder %>% group_by(continent) %>% filter (year==2007) %>% summarise(LifeExpectancy=mean(lifeExp)) %>% ggplot(aes(x=continent,y=LifeExpectancy, color=continent, fill=continent))+geom_bar(stat="identity")+scale_x_discrete(name="Continent")+scale_y_discrete(name="MeanLifeExpectancy")
```

###As seen in the previous graphs, we had focused only on one year, 2007. Now we’ve taken it one step further and wanted to look at all the years the data provided for each continent, side by side, to see the changes between each continent in each year as well as a trend as the years went on.

```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE, fig.width =10, fig.height=7}
gPop <- group_by(gapminder, year, continent)
summarise(gPop,Population=mean(pop))
gapminder %>% group_by(year, continent) %>%
  summarise(Population=mean(pop)) %>%
  ggplot(aes(x=year, y=Population, color=continent, fill=continent)) +
  geom_bar(stat="identity", position="dodge")+
  scale_x_discrete(name="Year", limits=c("1952", "1957", "1962",
                                         "1967", "1972", "1977",
                                         "1982", "1987", "1992",
                                         "1997", "2002", "2007"))

gGDP <- group_by(gapminder, year, continent)
summarise(gGDP, GDP=mean(gdpPercap))
gapminder %>% group_by(year, continent) %>%
  summarise(GDP=mean(gdpPercap)) %>%
  ggplot(aes(x=year, y=GDP, color=continent, fill=continent))+
  geom_bar(stat="identity", position="dodge")+
  scale_x_discrete(name="Year", limits=c("1952", "1957", "1962",
                                         "1967", "1972", "1977",
                                         "1982", "1987", "1992",
                                         "1997", "2002", "2007"))

gExp <- group_by(gapminder, year, continent)
summarise(gExp, LifeExpectancy=mean(lifeExp))
gapminder %>% group_by(year, continent) %>%
  summarise(LifeExpectancy=mean(lifeExp)) %>%
  ggplot(aes(x=year, y=LifeExpectancy, color=continent, fill=continent))+
  geom_bar(stat="identity", position="dodge")+
  scale_x_discrete(name="Year", limits=c("1952", "1957", "1962",
                                         "1967", "1972", "1977",
                                         "1982", "1987", "1992",
                                         "1997", "2002", "2007"))
```

###We used various graphs to examine trends. 

```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE, fig.width =7, fig.height=7}
ggplot(gapminder, aes(x=lifeExp, y=pop))+geom_point()+scale_y_log10()+geom_smooth(method="lm")

avg_gdp_continent <- gapminder %>% group_by(continent) %>% summarise(gdp=mean(gdpPercap)) %>%
   ggplot(aes(x=continent, y=gdp), ylab= "Average GDP")+geom_point()

avg_gdp_continent + ylab("Average GDP") + xlab("Continents")

mutate(gapminder,gdp=gdpPercap*pop) %>% filter(year==2007) %>%
   ggplot(aes(x=continent,y=gdp))+
   geom_point()+scale_y_log10() 
```

###These are graphs we were looking to see if a relationship existed. However, we realized that using the variables on their own gave very awkward distributions. 

```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE, fig.width =7, fig.height=7}
ggplot(gapminder, aes(x=pop, y=gdpPercap, color=continent))+geom_bar(stat="identity")+scale_x_discrete(name="Population")+scale_y_discrete(name="GDPperCapita")

ggplot(gapminder, aes(x=lifeExp, y=gdpPercap, color=continent))+geom_bar(stat="identity")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="GDPperCapita")

ggplot(gapminder, aes(x=lifeExp, y=year, color=continent))+geom_bar(stat="identity")+scale_x_discrete(name="LifeExpectancy")+scale_y_discrete(name="Year")
```



