---
title: "Assignment3"
author: "Kathleen Tran/Eric Almeida"
date: "3/6/2017"
output: pdf_document
---


```{r, echo=FALSE}
library(ggplot2)
library(curl)
library(dplyr)
library(tidyr)


gapminder_location <-"https://raw.githubusercontent.com/resbaz/r-novice-gapminder-files/master/data/gapminder-FiveYearData.csv"
gapminder<-read.csv(gapminder_location)

ggplot(subset(gapminder, continent %in% c("Asia", "Africa")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_point()+geom_smooth()+ggtitle("Looking at Life Expectancy via GDP per Capita in Asia and Africa")

ggplot(subset(gapminder, continent %in% c("Americas", "Europe")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_point()+ggtitle("Looking at Life Expectancy via GDP per Capita in the 
          Americas and Europe")

ggplot(subset(gapminder, continent %in% c("Oceania", "Asia")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_point()+ggtitle("Looking at Life Expectancy via GDP per Capita in Oceania and Asia")

ggplot(gapminder, aes(x=pop, y=gdpPercap, color=continent))+geom_bar(stat="identity")

ggplot(gapminder, aes(x=lifeExp, y=gdpPercap, color=continent))+geom_bar(stat="identity")

ggplot(gapminder, aes(x=lifeExp, y=year, color=continent))+geom_bar(stat="identity")

ggplot(subset(gapminder, continent %in% c("Asia", "Europe")),
       aes(x=lifeExp,
           y=gdpPercap,
           color=continent))+geom_bar(stat="identity")

ggplot(gapminder, aes(x=continent, color=continent, fill=continent))+geom_bar(stat="count")
```
```{r, echo=FALSE}
ggplot(gapminder, aes(x=year, y=gdpPercap, fill=continent))+geom_bar(stat="identity", position=position_dodge())

avg_gdp_continent <- gapminder %>% group_by(continent) %>% summarise(gdp=mean(gdpPercap)) %>%
  ggplot(aes(x=continent, y=gdp), ylab= "Average GDP")+geom_point() 

avg_gdp_continent + ylab("Average GDP") + xlab("Continents")

mutate(gapminder,gdp=gdpPercap*pop) %>% filter(year==2007) %>%
  ggplot(aes(x=continent,y=gdp))+
  geom_point()+scale_y_log10() 
```


As seen in the previous graphs, we had focused only on one year, 2007. Now we’ve taken it one step further and wanted to look at all the years the data provided for each continent, side by side, to see the changes between each continent in each year as well as a trend as the years went on.
```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE}

gGDP <- group_by(gapminder, year, continent)
summarise(gGDP, GDP=mean(gdpPercap))
gapminder %>% group_by(year, continent) %>%
  summarise(GDP=mean(gdpPercap)) %>%
  ggplot(aes(x=year, y=GDP, fill=continent))+
  geom_bar(stat="identity", position=position_dodge())+
  scale_x_discrete(name="Year", limits=c("1952", "1957", "1962",
                                         "1967", "1972", "1977",
                                         "1982", "1987", "1992",
                                         "1997", "2002", "2007"))

gPop <- group_by(gapminder, year, continent)
summarise(gPop,Population=mean(pop))
gapminder %>% group_by(year, continent) %>%
  summarise(Population=mean(pop)) %>%
  ggplot(aes(x=year, y=Population, color=continent, fill=continent)) +
  geom_bar(stat="identity", position="dodge")+
  scale_x_discrete(name="Year", limits=c("1952", "1957", "1962",
                                         "1967", "1972", "1977",
                                         "1982", "1987", "1992",
                                         "1997", "2002", "2007"))
